Data Exploration: IUCN and AquaMaps

### 2016-12-27 clip_fao/fao_clip_pac.Rmd
pushing recent updates to analysis and documentation based on reviewer comments
added environmental space analysis to fao_clip_pac.Rmd; now to put it into the manuscript.

@@

# am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
#                                    'v2016/int/am_cells_spp_prob0.csv')) 
# 
# am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
#   dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
#   mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
#   dplyr::select(-genus, -species) %>%
#   distinct()
# 
# hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
#   dplyr::select(loiczid, cellarea, oceanarea)
# 
# am_spp_cells2 <- hcaf %>%
#   left_join(am_spp_cells, by = 'loiczid')
# 
# 
# spp_area_df <- am_spp_cells2 %>%
#   group_by(am_sid) %>%
#   summarize(cellarea = sum(cellarea),
#             oceanarea = sum(oceanarea))
# 
# spp_area_df <- spp_area_df %>%
#   inner_join(am_spp_list, by = 'am_sid') %>%
#   mutate(area_per_cells = oceanarea/occurcells)
# 
# clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
#   filter(lim == 'e_lim') %>%
#   dplyr::select(am_sid)
# 
# spp_area_df <- spp_area_df %>%
#   mutate(log_area_cells = log(area_per_cells + 1),
#          clipped = am_sid %in% clipped_spp$am_sid)

@@

### compare ocean areas
summary(spp_area_df$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   #   1806    946600   2921000  10210000   7875000 501300000 
summary(spp_area_clipped$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 134900   4646000   7032000   8960000   9690000 111300000

### compare occurcells
summary(spp_area_df$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.00    6.00   16.00   57.13   49.00 7663.00 
summary(spp_area_clipped$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 3.00    6.00   15.00   26.86   33.00  494.00

### compare area_per_cells
summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #  190    50540   158000   485300   439100 44700000 
summary(spp_area_clipped$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 6578   202900   452400   827200   923300 26090000

>>>>

### how many of these are expert-reviewed?
clipped_spp_rev <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
  filter(reviewed == '1')

@@

# am_spp_cells <- read_csv(file.path(dir_anx, '_raw_data/aquamaps',
#                                    'd2015/csv/hcaf_sp_native_trunc.csv'))
# 
# am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
#   dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
#   mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
#   dplyr::select(-genus, -species) %>%
#   distinct()
# 
# hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
#   dplyr::select(loiczid, cellarea, oceanarea)
# 
# am_spp_cells2 <- hcaf %>%
#   left_join(am_spp_cells, by = 'loiczid')
# 
# spp_area_df <- am_spp_cells2 %>%
#   rename(am_sid = speciesid) %>%
#   group_by(am_sid) %>%
#   summarize(env_area  = sum(oceanarea * probability),
#             geom_area = sum(oceanarea))
# 
# spp_area_df <- spp_area_df %>%
#   inner_join(am_spp_list, by = 'am_sid') %>%
#   mutate(geom_area_per_cells = geom_area/occurcells,
#          env_area_per_cells  = env_area/occurcells)
# 
# clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
#   filter(lim == 'e_lim') %>%
#   dplyr::select(am_sid)
# 
# spp_area_df <- spp_area_df %>%
#   mutate(clipped = am_sid %in% clipped_spp$am_sid)

@@

### compare geometric space areas
summary(spp_area_df$geom_area)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   #   1806    946600   2921000  10210000   7875000 501300000 
summary(spp_area_clipped$geom_area)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 134900   4646000   7032000   8960000   9690000 111300000

### compare environmental space areas
summary(spp_area_df$env_area)
   # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   #   90    592200   1843000   6772000   5506000 402200000 
summary(spp_area_clipped$env_area)
   #  Min.  1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 75930  2994000   4909000   5989000   7156000  89480000

### compare occurcells by quadrant
summary(spp_area_df$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.00    6.00   16.00   57.13   49.00 7663.00 
summary(spp_area_clipped$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 3.00    6.00   15.00   26.86   33.00  494.00

### compare geometric area_per_cells
summary(spp_area_df %>% filter(occurcells != 0) %>% .$geom_area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #  190    50540   158000   485300   439100 44700000 
summary(spp_area_clipped$geom_area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 6578   202900   452400   827200   923300 26090000 
### compare environmental area_per_cells
summary(spp_area_df %>% filter(occurcells != 0) %>% .$env_area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #   30    32270   101600   303700   287400 30100000 
summary(spp_area_clipped$env_area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 3352   140000   311100   511300   619800 12480000

### sample species
### Rajella fyllae: Fis-131802
spp_area_df %>% filter(am_sid == 'Fis-131802') %>% select(geom_area, env_area, occurcells, geom_area_per_cells, env_area_per_cells)
    #   geom_area env_area occurcells geom_area_per_cells env_area_per_cells
    #       <dbl>    <dbl>      <int>               <dbl>              <dbl>
    # 1   5425916  3386906        116            46775.14           29197.47
### Ophiotrix plana: W-Oph-245456
spp_area_df %>% filter(am_sid == 'W-Oph-245456') %>% select(geom_area, env_area, occurcells, geom_area_per_cells, env_area_per_cells)
    #   geom_area env_area occurcells geom_area_per_cells env_area_per_cells
    #       <dbl>    <dbl>      <int>               <dbl>              <dbl>
    # 1   5426931  3053926          4             1356733           763481.5
